A Formalization of Complete Discrete Valuation Rings and Local Fields

### 2023-12-12 HAL/CPP_paper_v2.tex
reread every new modification from MI

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
%\setcopyright{acmcopyright}
\setcopyright{none} %% FOR ARXIV VERSION
%\copyrightyear{2023}
%\acmYear{2024}
%\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
%\acmConference[CPP 2024]{Certified Programs and Proofs 2024}{January 15--16, 2024}{London, UK}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
%\acmPrice{15.00}
%\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

@@

We make an exception to this rule when implementing mixed and equal characteristic local fields, since we do declare \code{valued} instances for them. This creates a mild diamond for $\padicCompl$ and $\FpXCompl$ but this does not cause trouble in our formalization, thanks to the comparison lemma \code{trivial_extension_eq_valuation}\href{https://github.com/LCFT-Lean/local_fields/blob/76ad487d09babdb0018f394a5634526637ee014a/src/discrete_valuation_ring/trivial_extension.lean#L58}{\extlink} mentioned above. Indeed, in all cases where a the inference system would create problems, the lemma allows us to disambiguate the situation: \fil{for an example, we urge the reader to inspect the proof of lemma} \code{FpX_int_completion.equiv_valuation_subring}\href{https://github.com/LCFT-Lean/local_fields/blob/76ad487d09babdb0018f394a5634526637ee014a/src/eq_characteristic/valuation.lean#L97}{\extlink}, \fil{proving that the unit ball inside $\FpXCompl$ as...}.

@@

Suppose that we had declared \code{discretely_normed_field} as an instance instead of as a definition, so that every discretely valued field
\begin{lstlisting}
(K : Type*) [field K] [hv : valued K $\wzmZZ$]
  [is_discrete hv.v]
\end{lstlisting}
\mi{would inherit a} \code{normed_field} \mi{instance.
Under this assumption, the typeclass inference system gets into an infinite loop when searching for an instance of
}
\code{
valuation_ring hv.v.valuation_subring.to_subring,}
\mi{because one of the first results that Lean tries to apply is}  \code{valuation_ring.of_field}, \mi{which reduces the problem to finding an instance of}
\code{field hv.v.valuation_subring.to_subring.}
\mi{This leads to a dead end (since the valuation subring is not a field); however, while trying to find this instance, the inference system finds} \code{normed_field.to_field}, \mi{and hence it starts to search for an instance of}
\code{normed_field hv.v.valuation_subring.to_subring}, \mi{for which it will try to apply the definition} \code{discretely_normed_field}, \mi{which as input requires a} \code{field} \mi{instance on the valuation subring. Hence, the system gets stuck in an infinite loop.}

However, observe that no problem arises when putting both a \code{valued} and a \code{normed_field} instance on a specific field, like $\QQ_p$\mi{, because the typeclass inference system is able to find these instances on the specific field, and it will not try to apply them to find instances in other types.}

>>>>

%%% The following is specific to CPP '24 and the paper
%%% 'A Formalization of Complete Discrete Valuation Rings and Local Fields'
%%% by María Inés de Frutos-Fernández and Filippo Alberto Edoardo Nuccio Mortarino Majno di Capriglio.
%%%
\setcopyright{acmlicensed}
\acmDOI{10.1145/3636501.3636942}
\acmYear{2024}
\copyrightyear{2024}
\acmSubmissionID{poplws24cppmain-p13-p}
\acmISBN{979-8-4007-0488-8/24/01}
\acmConference[CPP '24]{Proceedings of the 13th ACM SIGPLAN International Conference on Certified Programs and Proofs}{January 15--16, 2024}{London, UK}
\acmBooktitle{Proceedings of the 13th ACM SIGPLAN International Conference on Certified Programs and Proofs (CPP '24), January 15--16, 2024, London, UK}
\received{2023-09-19}
\received[accepted]{2023-11-25}

@@

We make an exception to this rule when implementing mixed and equal characteristic local fields, since we do declare \code{valued} instances for them. This creates a mild diamond for $\padicCompl$ and $\FpXCompl$ but this does not cause trouble in our formalization, thanks to the comparison lemma \code{trivial_extension_eq_valuation}\href{https://github.com/LCFT-Lean/local_fields/blob/76ad487d09babdb0018f394a5634526637ee014a/src/discrete_valuation_ring/trivial_extension.lean#L58}{\extlink} mentioned above. Indeed, in all cases where a the inference system would create problems, the lemma allows us to disambiguate the situation: \fil{for an example, the reader can inspect the proof of} \code{FpX_int_completion.equiv_valuation_subring}\href{https://github.com/LCFT-Lean/local_fields/blob/76ad487d09babdb0018f394a5634526637ee014a/src/eq_characteristic/valuation.lean#L97}{\extlink}, \fil{proving that the unit ball inside the normed field $\FpXCompl$ is isormorphic to the ring whose elements have $(X)$-adic valuation less than or equal to (\code{1:} $\wzmZZ$).}

@@

Suppose that we had declared \code{discretely_normed_field} as an instance instead of as a definition, so that every discretely valued field
\begin{lstlisting}
(K : Type*) [field K] [hv : valued K $\wzmZZ$]
  [is_discrete hv.v]
\end{lstlisting}
would \fil{automatically} inherit a \code{normed_field} instance.
Under this assumption, the typeclass inference system \fil{would} get into an infinite loop when searching for an instance of
%}
\code{valuation_ring hv.v.valuation_subring.to_subring,}\relax 
because one of the first results that Lean tries to apply is \code{valuation_ring.of_field}, \fil{that automatically infers a valuation ring structure on every field. In turn,} this reduces the problem \fil{at hand} to finding an instance of \code{field hv.v.valuation_subring.to_subring.}
This leads to a dead end (since the valuation subring is not a field); however, while trying to find this instance, the inference system finds \code{normed_field.to_field}\relax, and hence it starts to search for an instance of
\code{normed_field hv.v.valuation_subring.to_subring}, for which it will try to apply the definition \code{discretely_normed_field}, which as input requires a \code{field} instance on the valuation subring \fil{which recreates the problem discussed above, risulting in \st{. Hence, }}the system get\fil{\st{s}ting} stuck in an infinite loop.

However, observe that no problem arises when putting both a \code{valued} and a \code{normed_field} instance on a specific field, like $\QQ_p$, because the typeclass inference system is able to find these instances on the specific field, and it will not try to \fil{deduce\st{apply}} them \fil{\st{to find}from other} instances \fil{\st{in}on} other types.
